---
title: "Introduction to rrqueue"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rrqueue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Documenting things that work asynchronously is difficult.  This
document gives a tutorial-style overview of working with rrqueue.

# Getting started

The queue and workers can be started in any order, but it's easiest
to explain starting the queue first.

Suppose we have some simulation code; it needs to be in a file that
the queue can see.  For now, I'll use the file `myfuns.R` which is
the test code.  It has a function in it called `slowdouble` that
takes a number, sleeps for that many seconds, and then returns
twice the number.  It's useful for testing.

``` {r echo=FALSE, results="hide"}
rrqueue:::queue_clean(RedisAPI::hiredis(), "myqueue",
                      purge=TRUE, stop_workers=TRUE,
                      kill_local_workers=TRUE, wait_stop=0.1)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep="\n")
}
cpp_output <- function(x) lang_output(x, "c++")
r_output <- function(x) lang_output(x, "r")
yaml_output <- function(x) lang_output(x, "yaml")
plain_output <- function(x) lang_output(x, "plain")
```

You'll also need a running Redis server.  I have one operating with
the default parameters, so this works:
``` {r }
RedisAPI::hiredis()$PING()
```

Create queue called "myqueue", tell it to load the source file
"myfuns.R".  If it was to load packages, then passing
`packages=c("package1", "package2")` would indicate that workers
would need to load those packages, too.
``` {r }
obj <- rrqueue::queue("myqueue", sources="myfuns.R")
```

The message "creating new queue" here indicates that `rrqueue` did
not find any previous queues in place.  Queues are designed to be
re-attachable so we can immediately just do that:
``` {r }
obj <- rrqueue::queue("myqueue", sources="myfuns.R")
```

The message also notes that we have no workers available, so no
work is going to get done.  But we can still queue some tasks.

# Queuing tasks

The simplest sort of task queuing is to pass an expression into enqueue:
``` {r }
t <- obj$enqueue(1 + 1)
```

The expression is not evaluated but stored and will be evaluated on
the worker.  Saving the result of this gives a `task` object which
can be inspected.
``` {r }
t
```

The expression stored in the task:
``` {r }
t$expr()
```

The status of the task:
``` {r }
t$status()
```

The result of the task, which will throw an error if we try to call it:
``` {r error=TRUE}
t$result()
```

And how long the task has been waiting:
``` {r }
t$times()
```

Tasks can use local variables, too:
``` {r }
x <- 10
t2 <- obj$enqueue(x * 2)
t2$expr()
```

And because using unevaluated expressions can be problematic,
`rrqueue` has a standard-evaluation version (`enqueue_`) which takes
either strings representing expressions or quoted expressions:
``` {r }
obj$enqueue_(quote(x / 2))
```

Now we have three tasks:
``` {r }
obj$tasks_list()
```

All the tasks are waiting to be run:
``` {r }
obj$tasks_status()
```

We can get an overview of the tasks:
``` {r }
obj$tasks_overview()
```

# Starting workers

`rrqueue` includes a script `rrqueue_worker` for starting workers
from the command line (install with `rrqueue::install_scripts()`.
Workers can also be started from within R using the `worker_spawn`
function:
``` {r }
logfile <- tempfile()
wid <- rrqueue::worker_spawn("myqueue", logfile)
```
``` {r echo=FALSE}
Sys.sleep(.5)
```

This function returns the \emph{worker identifier}, which is also
printed to the screen.

It's probably informative at this point to read the logfile of the
worker to see what it did on startup:

``` {r results="asis", echo=FALSE}
plain_output(readLines(logfile))
```

The worker first prints a lot of diagnostic information to the
screen (or log file) indicating the name of the worker, the version
of rrqueue, machine information, and special keys in the database
where important information is stored.

Then after broadcasting that it is awake (`ALIVE`) it detected that
there was a controller on the queue and it attempts to construct
the environment that the controller wants `r paste("ENVIR", obj$envir_id)`.

After that, there are a series of `TASK_START`, `EXPR`, and
`TASK_COMPLETE` lines as each of the three tasks is processed.
``` {r }
obj$tasks_status()
```

The times here give an indication of the rrqueue overhead; the
running time of these simple expressions should be close to zero.
``` {r }
obj$tasks_times()
```

The task handle created before can now give a result:
``` {r }
t$result()
```

Similarly, results can be retrieved from the queue directly:
``` {r }
obj$task_result(1)
obj$task_result(2)
obj$task_result(3)
```

The worker that we created can be seen here:
``` {r }
obj$workers_list()
```

Queue a slower task; this time the `slowdouble` function.  This
will take 1s:
``` {r }
t <- obj$enqueue(slowdouble(1))
t$status()
Sys.sleep(.3)
t$status()
Sys.sleep(1)
t$status()
t$result()
```

Again, times are available:
``` {r }
t$times()
```

# Finishing up

Messaging will be properly dealt with in another vignette.
``` {r }
id <- obj$send_message("STOP")
obj$get_response(id, wid, wait=10)
```

``` {r results="asis", echo=FALSE}
plain_output(readLines(logfile))
```

worker is now in the exited list
``` {r }
obj$workers_list_exited()
```

The full log from our worker (dropping the first column which is
the worker id and takes up valuble space here):
``` {r }
obj$workers_log_tail(wid, Inf)[-1]
```
